---
title: "Probability Theory and Mathematical Statistics"
subtitle: "Script 3: Bernoulli trials"
author: "Dmitry Pavlyuk"
#date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Bernoulli formula
## Problem. 
Let a network consists of 8 computers. The probability of failure of each of them during a year is 10% (computers fail independently of one another).
What is the probability of:
* exactly one computer failure (during the same time frame)?
* no computer failures?
* more than one computer failure?

## Solution
```{r}
n = 8
p = 0.1
k = 1
# Manual calculation using the Bernoulli formula
p1f = choose(n,k)*p^k*(1-p)^(n-k)
```
Answer: probability of exactly one computer failure is `r p1f`

```{r}
# Calculation using the embedded function
p1 = dbinom(k, n,p)
```
Does the embedded *dbinom* function return the same result? `r all.equal(p1f,p1)`

```{r}
k = 0
p0 = dbinom(k, n,p)
p0
```
Answer: probability of no computer failures is `r p0`

```{r}
p_more1 = 1 - p0 - p1
```
Answer: probability of more than one computer failure is `r p_more1`

# Moivre-Laplace (normal) approximation
## Problem
Let the local network includes not 8 (as in the example in the previous section), but 800 computers. The probability of failure is 10% for each.
What is the probability that exactly 85 computer failures?
## Solution
```{r}
# Define a function for Moivre-Laplace (normal) approximation
approxMoivreLaplaceLocal = function(k, n, p){
	x = (k - n*p)/sqrt(n*p*(1-p))
	# Call the normal density function
  fi = dnorm(x)
	# Approximate calculation
	p = fi/sqrt(n*p*(1-p))
	return(p)
}

n = 800
p = 0.1
k = 85

p_ML = approxMoivreLaplaceLocal(k, n, p)
p_Bernoulli = choose(n,k)*p^k*(1-p)^(n-k)
```
Approximate probability (Moivre-Laplace approximation) = `r p_ML` 

Exact probability (Bernoulli formula) = `r p_Bernoulli` 

Compare calculated values: `r all.equal(p_ML,p_Bernoulli)`

## Problem
What is the probability that the number of failures will be from 80 to 100?

## Solution
```{r}
# Define a function for integral Moivre-Laplace (normal) approximation

approxMoivreLaplaceIntegral = function(k1, k2, n, p){
	x1 = (k1 - n*p)/sqrt(n*p*(1-p))
	x2 = (k2 - n*p)/sqrt(n*p*(1-p))
	Fi1 = pnorm(x1)
	Fi2 = pnorm(x2)
	p = Fi2-Fi1
	return (p)
}

k1 = 80
k2 = 100

p_ML = approxMoivreLaplaceIntegral(k1, k2, n, p)
p_Bernoulli = 0
for (k in k1:k2){
  p_Bernoulli = p_Bernoulli + choose(n,k)*p^k*(1-p)^(n-k)
}
```
Approximate probability (Moivre-Laplace approximation) = `r p_ML` 

Exact probability (Bernoulli formula) = `r p_Bernoulli` 

Compare calculated values: `r all.equal(p_ML,p_Bernoulli)`

# Poisson approximation

## Problem
Now let we now assume that probability of success is very small. For example, the probability that your computer will be damaged as a result of coffee spilled on it equal to 0.001.
Suppose you want to calculate probability of exactly one such computer failure of 800.

## Solution
```{r}
# Define a function for Poisson approximation
approxPoisson = function(k, n, p){
	if (p>0.5){
		k = n - k
		p = 1 - p
	}
	res = (n*p)^k/factorial(k)*exp(-n*p)
	return(res)
}

n = 800
p = 0.001
k = 1

p_Poisson = approxPoisson(k, n, p)
p_Bernoulli = choose(n,k)*p^k*(1-p)^(n-k)
```
Approximate probability (Poisson approximation) = `r p_Poisson` 

Exact probability (Bernoulli formula) = `r p_Bernoulli` 

Compare calculated values: `r all.equal(p_Poisson,p_Bernoulli)`


# Accuracy comparison of Poisson and Moive-Laplace approximation

```{r}
n = 50
k=seq(from=0, to=n)
p=0.1

# Calculating probabilities for all k values in [0,50] using the exact Bernoulli formula.
# Note the "magic" with vector calculation
pBin = choose(n,k)*p^k*(1-p)^(n-k)

# Calculating probabilities for all k values in [0,50] using the Moivre-Laplace approximation
pMoivreLaplace = approxMoivreLaplaceLocal (k,n,p)

# Calculating probabilities for all k values in [0,50] using the Poisson approximation
pPoisson = approxPoisson(k,n,p)

# Plot calculated probabilities

plot(pBin, type="l", ylab="Probability")
lines(pMoivreLaplace , type="l", col="red")
lines(pPoisson, type="l", col="blue")
```
```{r}
# Vectors of errors for different probability values
MoivreLaplaceErrors = c()
PoissonErrors = c()

# A cycle from 0 to 1 with a step 0.01
for (p in seq(0,1, by = 0.01)){
	pBin = dbinom(k,n,p)
	pMoivreLaplace = approxMoivreLaplaceLocal(k,n,p)
	pPoisson = approxPoisson(k,n,p)
	
	# Fill error vectors with relative squared errors
	MoivreLaplaceErrors = append(MoivreLaplaceErrors, sum(((pBin-pMoivreLaplace)^2)/pBin))
	PoissonErrors = append(PoissonErrors, sum(((pBin-pPoisson)^2)/pBin))
}

# Plot error values
plot(MoivreLaplaceErrors, type="l", xlab="Probability, %", ylab="Error")
lines(PoissonErrors, type="l", col="blue")
```

Make the conclusion about relative accuracy of Poisson and Moive-Laplace approximations!