---
title: "Probability Theory and Mathematical Statistics"
subtitle: "Practise 5"
author: "Ivans Gercevs"
#: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---
GitHub repository https://github.com/foxybbb/TSI_Probability-Statistics


# Create a new R markdown and include the codesfor the central limit theorem illustrationfrom learning-by-doingmaterials 08


## Central Limit Theorem

#_Convergence of a sum of random variables to the normal distribution_

# Analyse convergence in distribution, provided by the central limit theorem, by plotting the empirical cumulative distribution function (ecdfcommand) for different distribution (exponential and uniform) and number of summands.

### Sum of uniformly distributed rabdon variables
```{r  figures-side, fig.show="hold", out.width="50%"}
for(i in c(2, 4, 30, 300)) {
	# Repeat the experiment 10000 times
	sums= replicate(10000,sum(runif(i)))
	
	# Sample mean and standard deviation
	mu=mean(sums)
	s = sd(sums)
	
	# Plot the histogram of sums
	plot(ecdf((sums-mu)/s),main=paste("Function  of", i," uniformly empirical cumulative distribution variables"))
	hist((sums-mu)/s,freq=FALSE, main=paste("Sums of", i,"uniformly distributed variables"))
	
	# Add the normal curve for comparison
	x <- seq(0,i,length=100)
	curve(dnorm(x), col = "red", add = TRUE)
}

```


## Sum of exponentially distributed rabdon variables
```{r  fig.show="hold", out.width="50%"}
for(i in c(2, 4, 30, 300)) {
  sums = replicate(10000, sum(rexp(i)))
  mu = mean(sums)
  s = sd(sums)
  plot(
    ecdf((sums - mu) / s),
    main = paste(
      "Function  of",
      i,
      "exponentially empirical cumulative distribution variables"
    )
  )
  hist((sums - mu) / s,
       freq = FALSE,
       main = paste("Sums of", i, "exponentially distributed variables")
  )
  x <- seq(0, i, length = 100)
  curve(dnorm(x), col = "red", add = TRUE)
  
}

```




# Implement a function for calculating the approximation accuracy metric as:

```{r}
# function which calculating ...
kolomogorov_smirnov_mthd <- function(sums) {
  x <- seq(0, i, length = length(sums))
  # Sample mean and standard deviation
  mu = mean(sums)
  s = sd(sums)
  # Create normal distribution  using mean, sd
  norm = dnorm(x, mean = mu, sd = s)
  return (max(abs(norm - sums)))
  
}
```
```{r}
pr1 = c()
pr2 = c()

# testing with function using uniform distributed variables
for (i in c(2, 4, 30, 300)) {
  sums = replicate(10000, sum(runif(i)))
  pr1 <- append (pr1, kolomogorov_smirnov_mthd(sums))
}
# testing with function using exponentially distributed variables
for (i in c(2, 4, 30, 300)) {
  sums = replicate(10000, sum(rexp(i)))
  pr2 <- append (pr2, kolomogorov_smirnov_mthd(sums))
}
```

For uniform distributed variables `r pr1`


For exponentially distributed variables `r pr2`

# Develop a function St(n, k, A)that returns a random sample of size k=1000 for the following statistic (distribution of summands can be chosen arbitrary)
```{r}

St <- function(n, k = 1000, A) {
  sums = replicate(k, sum(rexp(2)))
  mu = mean(sums)
  s = sd(sums)
  
  return (mu - A) / (s / sqrt(n))
}

print(St(10,1000,2))
```
# Calculate empirical 90%, 95%, 99% quantiles for the constructed statistic for n=10 and compare them with sampling distributions you know. Make the conclusion and include it into the markdown
```{r, echo=FALSE}
# В пррцессе выполнения было не явно выявленно условие реалиации конечного ответа, поэтому для тестирования, будет выбран такой вариант
x = sums

```
```{r}

quantile(x,0.90)
quantile(x,0.95)
quantile(x,0.99)

```
# Plot empirical 95% quantile values versus n; make the conclusionand include it into the markdown.
```{r}

 print(qnorm(.95, mean(x), sd(x)))

```

