---
title: "Probability Theory and Mathematical Statistics"
subtitle: "Practise 5"
author: "Ivans Gercevs"
#: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---
GitHub repository https://github.com/foxybbb/TSI_Probability-Statistics


# Create a new R markdown and include the codesfor the central limit theorem illustrationfrom learning-by-doingmaterials 08


## Central Limit Theorem

#_Convergence of a sum of random variables to the normal distribution_

### Sum of uniformly distributed rabdon variables
```{r  figures-side, fig.show="hold", out.width="50%"}
for(i in c(2, 4, 30, 300)) {
	# Repeat the experiment 10000 times
	sums= replicate(10000,sum(runif(i)))
	
	# Sample mean and standard deviation
	mu=mean(sums)
	s = sd(sums)
	
	# Plot the histogram of sums
	hist((sums-mu)/s,freq=FALSE, main=paste("Sums of", i,"uniformly distributed variables"))
	
	# Add the normal curve for comparison
	x <- seq(0,i,length=100)
	curve(dnorm(x), col = "red", add = TRUE)
}
```


## Sum of exponentially distributed rabdon variables
```{r  fig.show="hold", out.width="50%"}
for(i in c(2, 4, 30, 300)) {
	sums= replicate(10000,sum(rexp(i)))
	mu = mean(sums)
	s = sd(sums)
	hist((sums-mu)/s,freq=FALSE, main=paste("Sums of", i,"exponentially distributed variables"))
	x <- seq(0,i,length=100)
	curve(dnorm(x), col = "red", add = TRUE)
}
```

# Analyse convergence in distribution, provided by the central limit theorem, by plotting the empirical cumulative distribution function (ecdfcommand) for different distribution (exponential and uniform) and number of summands.

```{r}
x = seq(0,100,by = 0.01)




exponent = exp(x)
pr1 = ecdf(exponent)
plot(exponent,type="l")

z = seq(-3, 3, by=0.01)
pr1 = P(z) 

plot(pr1,type="l")


unif = dunif(x,min = 15,max = 50)
pr2 = ecdf(unif)
plot(unif,type="l")
plot(pr2)

rand = rnorm(x)
pr2 = ecdf(rand)
plot(rand)
plot(pr2)


```



# Implement a function for calculating the approximation accuracy metric as:

```{r}
monte_carlov_mthd <- function(x) {
  kd= seq(from = 0,to = x)
  pr = c()
  for(i in kd){
  #  pr <- append(abs(Fnormal(i)-Fsum(i))
    
  }
  return (max(pr))
}
```

```{r}
St <- function(n, k, A){
  #(k-A)/(s/sqrt(n))
  return ()
}
```