---
title: "Probability Theory and Mathematical Statistics"
subtitle: "Script 8: Law of Large Numbers"
author: "Dmitry Pavlyuk"
#date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# Bernoulli's law of large numbers 

_Convergence of the relative frequency to the theoretical probability_
```{r}
# Define number of random experiments (n) and number of positive outcomes (k)
p0 = 1/6
n = 0
k = 0

# Define a vector for storing intermediate relative frequencies
relFrequencies = c()

# Define number of simulations
N = 3000

# Simulate the experiment N times
for (i in 1:N){
	# result - simple simulation of a die
	result = sample(c(1,2,3,4,5,6), 1)
	
	# increase number of conducted experiments
	n = n + 1

	# increase number of positive outcomes (if '6' is obtained)
	if (result == 6) k = k + 1

	# Calculate the intermedirelative frequency value and add it to the vector
	w = k/n
	relFrequencies = append(relFrequencies, w)

}

# Calculate 95% interval using the estimate of Bernoulli's theorem
gamma = 0.95
err = sqrt(relFrequencies*(1-relFrequencies)/((1:N)*gamma))
```

Plot relative frequencies and confidence interval 
against number of conducted experiments
```{r}
	plot(relFrequencies, type="l")
	lines(1:N, relFrequencies-err, lty = 'dashed', col = 'blue')
  lines(1:N, relFrequencies+err, lty = 'dashed', col = 'blue')
	abline(h=p0, col="red")
```

Print out first 10 values of relative frequencies
```{r}
kable(head(relFrequencies, n=10))
```

Print out last 10 values of relative frequencies
```{r}
kable(tail(relFrequencies, n=10))
```



# Chebyshev's law of large numbers 

_Convergence of the mean to the theoretical expected value_

```{r}
# Define a vector for storing intermediate mean values
sample.means = c()

# Theoretical expected value
mu = 10

# Theoretical standard deviation
sd = 10

# Intermediate sum of values
sumX = 0

# Simulate the experiment N times
for (i in 1:N){
	# result - implementation of the normally distributed random variables
	result = rnorm(1,mean=mu, sd=sd)
	
	# Add newly generated value to the total sum
	sumX = sumX + result
	
	# Intermediate mean value
	m = sumX/i

	# Add intermediate mean values to the vector
	sample.means = append(sample.means, m)
}

# Calculate 95% interval using the estimate of Chebyshev's theorem
gamma = 0.95
err = sd / sqrt((1:N)*gamma)
```


Plot mean values and related confidence interval 
against number of conducted experiments
```{r}
	plot(sample.means, type="l")
	lines(1:N, sample.means-err, lty = 'dashed', col = 'blue')
  lines(1:N, sample.means+err, lty = 'dashed', col = 'blue')
	abline(h=mu, col="red")
```

Print out first 10 values of the mean
```{r}
kable(head(sample.means, n=10))
```

Print out last 10 values of the mean
```{r}
kable(tail(sample.means, n=10))
```


