---
title: "Probability Theory and Mathematical Statistics"
subtitle: "Script 9: Sample Statistics"
author: "Dmitry Pavlyuk"
##date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.height=4) 
library(knitr)
```


# Sample statistics

## Loading data
```{r}
# load data from CSV file
flats <- read.csv("data.csv", header=T, sep=";")
kable(head(flats), caption ="First observations")
n = nrow(flats)
```
Sample volume = `r n`

## Descriptive statistics for quantitative data
```{r}

kable(summary(flats), caption ="Summary statistics")

# Define a function for the sample mode calculation
mode <- function(x) {
  ux <- unique(x)
  return(ux[which.max(tabulate(match(x, ux)))])
}
```


### Descriptive statistics for Rooms
```{r}
df = data.frame(unclass(summary(flats$Rooms)))
colnames(df)<-c("Rooms")
kable(df,caption ="5-number statistics (Rooms)")
```
Sample mode = `r mode(flats$Rooms)`
```{r}
kable(table(flats$Rooms),caption ="Absolute frequencies (Rooms)")
wj = table(flats$Rooms)/n
kable(round(wj, digits=3),caption ="Relative frequencies (Rooms)")
plot(wj, type="b", main="Distribution (Rooms)")

plot(ecdf(flats$Rooms), main="Empirical cumulative distribution function (Rooms)")
```

### Descriptive statistics for Price
```{r}
df = data.frame(unclass(summary(flats$Price)))
colnames(df)<-c("Price")
kable(df,caption ="5-number statistics (Price)")
```
Sample mode = `r mode(flats$Price)`
```{r}
boxplot(flats$Price, main="Box-and-whisker plot (Price)")

# Statistical distribution
wj = hist(flats$Price, plot=FALSE)
df= data.frame(breaks = wj$breaks[-1],AbsFreq=wj$counts,RelFreq=wj$counts/n)
kable(df, caption ="Statistical distribution series (Price)")
hist(flats$Price, main="Histogram (Price)")

plot(ecdf(flats$Price), main="Empirical cumulative distribution function (Price)")
```

### Descriptive statistics for Area
```{r}
df = data.frame(unclass(summary(flats$Area)))
colnames(df)<-c("Area")
kable(df,caption ="5-number statistics (Area)")
```
Sample mode = `r mode(flats$Area)`
```{r}
boxplot(flats$Area, main="Box-and-whisker plot (Area)")


# Statistical distribution
wj = hist(flats$Area, plot=FALSE)
df= data.frame(breaks = wj$breaks[-1],AbsFreq=wj$counts,RelFreq=wj$counts/n)
kable(df, caption ="Statistical distribution series (Area)")
hist(flats$Area, main="Histogram (Area)")

plot(ecdf(flats$Area), main="Empirical cumulative distribution function (Area)")
```


## Descriptive statistics for quanlitative data
```{r}
library(qcc)
```

### Descriptive statistics for Type
Sample mode = `r mode(flats$Type)`
```{r}
kable(table(flats$Type),caption ="Absolute frequencies (Type)")
wj = table(flats$Type)/n
kable(round(wj, digits=3),caption ="Relative frequencies (Type)")
barplot(wj/n, main="Bar plot (Type")
pareto.chart(wj/n, main="Pareto Chart (Type)")

```
### Descriptive statistics for District
Sample mode = `r mode(flats$District)`
```{r}
kable(table(flats$District),caption ="Absolute frequencies (District)")
wj = table(flats$District)/n
kable(round(wj, digits=3),caption ="Relative frequencies (District)")
barplot(wj/n, main="Bar plot (District)")
pareto.chart(wj/n, main="Pareto Chart (District)")

```

### Two-variable statistics

Contigency table for Type and District

```{r}
ct = table(flats$District, flats$Type)
kable(ct, caption ="Contigency table - absolute frequencies (Type, District)")

kable(margin.table(ct, 1), caption ="Marginal frequencies (Type)")
kable(margin.table(ct, 2), caption ="Marginal frequencies (District)")

kable(prop.table(ct), caption ="Contingency table - relative frequencies (to sample size)")
kable(prop.table(ct,1), caption ="Contingency table - relative frequencies (to row sums)")
kable(prop.table(ct,2), caption ="Contingency table - relative frequencies (to column sums)")
```
